---
title: "backup"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r, eval=FALSE, echo=FALSE}
# p <- ggplot(df.covid %>% filter(GDPCAP>0), aes(x=GDPCAP), ) +
#   geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.8)
# p
```

```{r}
attach(df.covid)
par(mfrow=c(2,2))
hist(POP, main="histogram of ")
hist(GDPCAP, main="Scatterplot of wt vs. mpg")
hist(HDI, main="Histogram of wt")
hist(TC, main="Boxplot of wt")
```

HDI
```{r}
p5 <- df.covid %>%
  filter( HDI>0 ) %>%
  ggplot( aes(x=HDI)) +
    geom_histogram( binwidth=0.1, fill="#42e0f5", color="#e9ecef", alpha=0.9) +
    ggtitle("HDI histogram") +
    theme(
      plot.title = element_text(size=20), panel.background = element_blank(), axis.line = element_line(colour = "grey")
    )
p

#p <- ggplot(df.covid %>% filter(GDPCAP>0), aes(x=GDPCAP)) + 
#  geom_histogram(binwidth=0.45, fill="#69b3a2", color="#e9ecef", alpha=0.9)
#p
```

GPDCAP

```{r}
p6 <- df.lastdate %>%
  filter( GDPCAP>0 ) %>%
  ggplot( aes(x=GDPCAP)) +
    geom_histogram( binwidth=0.1, fill="#42e0f5", color="#e9ecef", alpha=0.9) +
    ggtitle("GDPCAP histogram") +
    theme(
      plot.title = element_text(size=20), panel.background = element_blank(), axis.line = element_line(colour = "grey")
    )
p

#p <- ggplot(df.covid %>% filter(GDPCAP>0), aes(x=GDPCAP)) + 
#  geom_histogram(binwidth=0.45, fill="#69b3a2", color="#e9ecef", alpha=0.9)
#p
```
This histogram shows that the distribution of GDP per Capita (GDPCAP) is left-skewed

```{r}
p <- df.covid %>% filter(GDPCAP>0)
p$GDPCAP %>% median()
```
The median value of data is 9.54




Total Cases
```{r}
p4 <- df.lastdate %>%
  filter( TC>0 ) %>%
  ggplot( aes(x=TC)) +
    geom_histogram( binwidth=0.1, fill="#42e0f5", color="#e9ecef", alpha=0.9) +
    ggtitle("Total Cases histogram") +
    theme(
      plot.title = element_text(size=20), panel.background = element_blank(), axis.line = element_line(colour = "grey")
    )
p4

#p <- ggplot(df.covid %>% filter(GDPCAP>0), aes(x=GDPCAP)) + 
#  geom_histogram(binwidth=0.45, fill="#69b3a2", color="#e9ecef", alpha=0.9)
#p
```


Total Death
```{r}
p3 <- df.covid %>%
  filter( TD>0 ) %>%
  ggplot( aes(x=TD)) +
    geom_histogram( binwidth=0.1, fill="#42e0f5", color="#e9ecef", alpha=0.9) +
    ggtitle("Total Death histogram") +
    theme(
      plot.title = element_text(size=20), panel.background = element_blank(), axis.line = element_line(colour = "grey")
    )
p

#p <- ggplot(df.covid %>% filter(GDPCAP>0), aes(x=GDPCAP)) + 
#  geom_histogram(binwidth=0.45, fill="#69b3a2", color="#e9ecef", alpha=0.9)
#p
```

Sexually Transmitted Infection
```{r}
p2 <- df.covid %>%
  filter( STI>0 ) %>%
  ggplot( aes(x=STI)) +
    geom_histogram( binwidth=0.1, fill="#42e0f5", color="#e9ecef", alpha=0.9) +
    ggtitle("Sexually Transmitted Infection histogram") +
    theme(
      plot.title = element_text(size=20), panel.background = element_blank(), axis.line = element_line(colour = "grey")
    )
p

#p <- ggplot(df.covid %>% filter(GDPCAP>0), aes(x=GDPCAP)) + 
#  geom_histogram(binwidth=0.45, fill="#69b3a2", color="#e9ecef", alpha=0.9)
#p
```


Population
```{r}
p <- df.covid %>%
  filter( POP>0 ) %>%
  ggplot( aes(x=POP)) +
    geom_histogram( binwidth=0.1, fill="#42e0f5", color="#e9ecef", alpha=0.9) +
    ggtitle("Population histogram") +
    theme(
      plot.title = element_text(size=20), panel.background = element_blank(), axis.line = element_line(colour = "grey")
    )
p

#p <- ggplot(df.covid %>% filter(GDPCAP>0), aes(x=GDPCAP)) + 
#  geom_histogram(binwidth=0.45, fill="#69b3a2", color="#e9ecef", alpha=0.9)
#p
```



```{r}
p <- ggplot(df.covid %>% filter(HDI>0), aes(x=HDI)) + 
  geom_histogram(binwidth=0.02, fill="#69b3a2", color="#e9ecef", alpha=0.9)
p
```


```{r}
p <- df.covid %>%
  filter( HDI>0 ) %>%
  ggplot( aes(x=HDI)) +
    geom_histogram( binwidth=0.01, fill="#42e0f5", color="#e9ecef", alpha=0.9) +
    ggtitle("Bin size = 0.01") +
    theme(
      plot.title = element_text(size=20), panel.background = element_blank(), axis.line = element_line(colour = "grey")
    )
p
```



```{r}
# Stacked density plot:
p <- ggplot(data=df.covid, aes(x=price, group=cut, fill=cut)) +
    geom_density(adjust=1.5, position="fill")
```


I erased the HDI = 0 part since it is no use in the research.
```{r}
# p <- ggplot(df.covid, aes(x=TC)) + 
#   geom_histogram(binwidth=0.2, fill="#69b3a2", color="#e9ecef", alpha=0.9)
# p
```
 
```{r}
# p <- ggplot(df.covid, aes(x=TD)) + 
#   geom_histogram(binwidth=0.2, fill="#69b3a2", color="#e9ecef", alpha=0.9)
# p
```

```{r}
# p <- ggplot(df.covid, aes(x=STI)) + 
#   geom_histogram(binwidth=0.88, fill="#69b3a2", color="#e9ecef", alpha=0.9)
# p
```

```{r}
# p <- ggplot(df.covid %>% filter(POP>0), aes(x=POP)) + 
#   geom_histogram(binwidth=0.5, fill="#69b3a2", color="#e9ecef", alpha=0.9)
# p
```

```{r}
df.covid2 <- df.covid%>%filter((COUNTRY %in% c("China", "United States")))

df.covid2 <- df.covid2 %>% filter(GDPCAP>0)

fit = lm(TC ~ 1 + GDPCAP, data = df.covid2)

fit%>%summary()

p <- ggplot(df.covid2, aes(x=DATE, y=TC, color = "#42e0f5")) +
      geom_point(alpha = 0.3, size = 0.1) +
      geom_smooth(method = "lm", se = T, color = "blue", fill="black", size = 1) +
  theme_bw()
p
```

```{r}
usa<-df.covid%>%filter(COUNTRY=="United States")
chn<-df.covid%>%filter(COUNTRY=="China")
```

```{r}
t.test(usa$TC,chn$TC)
```

```{r}
df.covid%>%filter(COUNTRY=="United States"|COUNTRY=="China")%>%
  ggplot(.,aes(x=COUNTRY, y=TD))+
  geom_bar(stat = "identity")
```


```{r}
# Set relative size of marginal plots (main plot 10x bigger than marginals)
#p2 <- ggMarginal(p, type="histogram", size=5)

# Set relative size of marginal plots (main plot 10x bigger than marginals)
#p1 <- ggMarginal(p, type="histogram", size=5)
 
# Custom marginal plots:
#p2 <- ggMarginal(p, type="histogram", fill = "slateblue", xparams = list(  bins=10))
 
# Show only marginal plot for x axis
#p3 <- ggMarginal(p, margins = 'x', color="purple", size=4)
```
 


\```{r}
 df.covid <- df.covid%>%filter(TC>0, GDPCAP>0)
# classic plot :
p <- ggplot(df.covid, aes(x=GDPCAP, y=TC, color = "#42e0f5", size=1)) +
      geom_point() +
      theme(legend.position="none")
 
# Set relative size of marginal plots (main plot 10x bigger than marginals)
p1 <- ggMarginal(p, type="histogram", size=5)
p1

```

```{r}

```


RELATION BETWEEN GDPCAP AND TC

```{r}
df.covid1 <- df.covid%>% filter(GDPCAP>0)
p <- ggplot(df.covid1, aes(x=GDPCAP, y=TC, color = "#42e0f5", size=1)) +
      geom_point() +
      theme(legend.position="none")
p
```

```{r}
p <- ggMarginal(p, type="histogram", size=5)
p
```

```{r}
cor.test(df.covid$TC, df.covid$GDPCAP)
```

```{r}
fit = lm(TC ~ 1 + GDPCAP, data = df.covid1)

fit%>%summary()
```



RELATION BETWEEN GDPCAP AND TC

```{r}

```




RELATION BETWEEN POP AND GDPCAP

```{r}
df.covidgdp <- df.covid %>% filter(GDPCAP>0)

fit = lm(TC ~ 1 + GDPCAP, data = df.covidgdp)

fit%>%summary()

p <- ggplot(df.covidgdp, aes(x=POP, y=GDPCAP, color = "#42e0f5")) +
      geom_point(alpha = 0.3, size = 0.1) +
      geom_smooth(method = "lm", se = T, color = "blue", fill="black", size = 1) +
  theme_bw()
p
```


RELATION BETWEEN GDPCAP AND TD

```{r}
df.covidgf <- df.covid %>% filter(GDPCAP>0)

fit = lm(TD ~ 1 + GDPCAP, data = df.covidgf)

fit%>%summary()

p <- ggplot(df.covidgf, aes(x=GDPCAP, y=TD, color = "#42e0f5")) +
      geom_point(alpha = 0.3, size = 0.1) +
      geom_smooth(method = "lm", se = T, color = "blue", fill="black", size = 1) +
      theme_bw()
p
```



RELATION BETWEEN POP AND TC

```{r}
p <- ggplot(df.covid, aes(x=POP, y=TC, color = "#42e0f5", size=1)) +
      geom_point() +
      theme(legend.position="none")
 
# Set relative size of marginal plots (main plot 10x bigger than marginals)
p1 <- ggMarginal(p, type="histogram", size=5)
 
# Custom marginal plots:
#p2 <- ggMarginal(p, type="histogram", fill = "slateblue", xparams = list(  bins=10))
 
# Show only marginal plot for x axis
#p3 <- ggMarginal(p, margins = 'x', color="purple", size=4)

p1
```

```{r}

```

```{r}
fit = lm(TC ~ 1 + POP, data = df.covid)

fit%>%summary()
```


RELATION BETWEEN STI AND TC

```{r}
df.covidsf <- df.covid%>% filter(STI>0)
```


```{r}
p <- ggplot(df.covidsf, aes(x=STI, y=TC, color = "#42e0f5", size=1)) +
      geom_point() +
      theme(legend.position="none")
 
# Set relative size of marginal plots (main plot 10x bigger than marginals)
p1 <- ggMarginal(p, type="histogram", size=5)
 
# Custom marginal plots:
#p2 <- ggMarginal(p, type="histogram", fill = "slateblue", xparams = list(  bins=10))
 
# Show only marginal plot for x axis
#p3 <- ggMarginal(p, margins = 'x', color="purple", size=4)

p1
```

```{r}
cor.test(df.covid$STI, df.covid$TC)
```

```{r}
fit = lm(TC ~ 1 + STI, data = df.covid)

fit%>%summary()
```

```{r}
fit = lm(TC ~ 1 +STI, data = df.covidsf)

fit%>%summary()

p <- ggplot(df.covid, aes(x=STI, y=TC, color = "#42e0f5")) +
      geom_point(alpha = 0.3, size = 0.1) +
      geom_smooth(method = "lm", se = T, color = "blue", fill="black", size = 1) +
  theme_bw()
p
```


RELATION BETWEEN STI AND GDPCAP
```{r}
df.covidsgf <- df.covid%>% filter(GDPCAP>0)
```


```{r}
p <- ggplot(df.covidsgf, aes(x=STI, y=GDPCAP, color = "#42e0f5", size=1)) +
      geom_point() +
      theme(legend.position="none")
 
# Set relative size of marginal plots (main plot 10x bigger than marginals)
p1 <- ggMarginal(p, type="histogram", size=5)
 
# Custom marginal plots:
#p2 <- ggMarginal(p, type="histogram", fill = "slateblue", xparams = list(  bins=10))
 
# Show only marginal plot for x axis
#p3 <- ggMarginal(p, margins = 'x', color="purple", size=4)

p1
```

```{r}
cor.test(df.covid$STI, df.covid$GDPCAP)
```

```{r}
fit = lm(TC ~ 1 + STI, data = df.covid)

fit%>%summary()
```

```{r}
fit = lm(TC ~ 1 + STI, data = df.covidsgf)

fit%>%summary()

p <- ggplot(df.covidsgf, aes(x=STI, y=TC, color = "#42e0f5")) +
      geom_point(alpha = 0.3, size = 0.1) +
      geom_smooth(method = "lm", se = T, color = "blue", fill="black", size = 1) +
  theme_bw()
p
```



```{r}
fit = lm(TC ~ 1 + GDPCAP, data = df.covidsgf)

fit%>%summary()

p <- ggplot(df.covidsgf, aes(x=STI, y=GDPCAP, color = "#42e0f5")) +
      geom_point(alpha = 0.3, size = 0.1) +
      geom_smooth(method = "lm", se = T, color = "blue", fill="black", size = 1) +
  theme_bw()
p
```

```{r}
df.covidstf <- df.covid%>% filter(STI>0)
```


```{r}
p <- ggplot(df.covidstf, aes(x=STI, y=TD, color = "#42e0f5", size=1)) +
      geom_point() +
      theme(legend.position="none")
 
# Set relative size of marginal plots (main plot 10x bigger than marginals)
p1 <- ggMarginal(p, type="histogram", size=5)
 
# Custom marginal plots:
#p2 <- ggMarginal(p, type="histogram", fill = "slateblue", xparams = list(  bins=10))
 
# Show only marginal plot for x axis
#p3 <- ggMarginal(p, margins = 'x', color="purple", size=4)

p1
```

```{r}
cor.test(df.covid$STI, df.covid$TD)
```

```{r}
fit = lm(TC ~ 1 + GDPCAP, data = df.covidsgf)

fit%>%summary()

p <- ggplot(df.covidsgf, aes(x=STI, y=TD, color = "#42e0f5")) +
      geom_point(alpha = 0.3, size = 0.1) +
      geom_smooth(method = "lm", se = T, color = "blue", fill="black", size = 1) +
  theme_bw()
p
```


```{r}
fit = lm(TC/POP ~ 1 + GDPCAP+HDI+STI, data = df.lastdate)

fit%>%summary()
```

```{r}
#head(df.covid)

# classic plot :
p1 <- ggplot(df.lastdate, aes(x=GDP_norm, y=HDI_norm, color = "#42e0f5", size=1)) +
      geom_point(alpha = 1, size = 1) +
      geom_smooth(method = "lm", se = T, color = "blue", fill="black", size = 1) +
  theme_classic()+
      theme(legend.position ="none")
  
 



# classic plot :
p2 <- ggplot(df.lastdate, aes(x=TC_norm, y=TD_norm, color = "#42e0f5", size=1)) +
      geom_point(alpha = 1, size = 1) +
      geom_smooth(method = "lm", se = T, color = "blue", fill="black", size = 1) +
  theme_classic()+
      theme(legend.position ="none")



p3 <- ggplot(df.lastdate, aes(x=GDP_norm, y=TC_norm, color = "#42e0f5")) +
      geom_point(alpha = 1, size = 1) +
      geom_smooth(method = "lm", se = T, color = "blue", fill="black", size = 1) +
  theme_classic()+
      theme(legend.position = "none")

p4 <- ggplot(df.lastdate, aes(x=GDP_norm, y=TD_norm, color = "#42e0f5")) +
      geom_point(alpha = 1, size = 1) +
      geom_smooth(method = "lm", se = T, color = "blue", fill="black", size = 1) +
  theme_classic()+
      theme(legend.position = "none")


p1 + p2 + p3 +p4
  plot_layout(ncol=1) &
  theme(
      plot.title = element_text(size=20), panel.background = element_blank(), axis.line = element_line(colour = "grey")
     ) 
```

```{r}
dus <- df.covid %>% filter(COUNTRY == "China")
d1 <- diff(dus$TC)
x2 <- dus$DATE[1:length(dus$DATE)-1]
d2 <- data.frame(d1)
dp1 <- ggplot(d2, aes(x=x2 ,y=d1)) +
  geom_line() + 
  xlab("")
dp1
```

```{r}
dus <- df.covid %>% filter(COUNTRY == "United States")
d1 <- diff(dus$TC)
x2 <- dus$DATE[1:length(dus$DATE)-1]
d2 <- data.frame(d1)
dp1 <- ggplot(d2, aes(x=x2 ,y=d1)) +
  geom_line() + 
  xlab("")
dp1
```